Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-02-23 SIR_manuscript.tex
introduction was updated

Because individuals who have been tested or received a positive test are likely to self-isolate---reducing or eliminating  their potentially infectious contacts---the dynamics of an epidemic will depend on who gets tested and
why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
the more strongly it is focused on people who are infectious.
This level of focus depends in turn on the purpose and design of testing programs.
(1) When testing is done for the purposes of disease surveillance \citep{foddai2020base}
we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref?}.

@@

In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic model with epidemic and testing components. Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$, which is useful to estimate the early dynamics of an epidemic. This index is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}. 
In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$. 
To control the spread of disease in the population, \david{the effective} $\Rnum$ must be $<1$.
Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying testing processes.
\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{edited, but still needs work}

>>>>

Because individuals who have been tested or received a positive test are likely to self-isolate---reducing or eliminating  their potentially infectious contacts---the dynamics of an epidemic will depend on who gets tested and
why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
the more strongly it is focused on people who are infectious.
This level of focus depends in turn on the purpose and design of testing programs. 
When testing is done for the purposes of disease surveillance \citep{foddai2020base}
we might expect tests to be assigned randomly across the population, thus we characterize it as \emph{random} testing strategy. In an ideal world, some bias in selection for testing is probably inevitable, although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.

@@

The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susciptable population, before the disease becomes extinct. This is determined by studying the basic reproduction number $\Rnum$. It is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}. 
In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$, and to decline over time when $\Rnum<1$. 
Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.

In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic SIR-type model with epidemic and testing components. Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on the epidemic dynamics. 
Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying epidemic and testing parameters.
\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{major changes done}
